Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 conclusion.tex
Added conclusion.

Parallel run-time monitoring techniques are attractive solutions for improving
the reliability, security, and debugging capabilities of systems. In this
paper, we have presented an architecture that reduces the overheads of
monitoring through filtering and adjustable partial monitoring. This is done by
using a hardware dataflow tracking engine in order to track and filter out
monitoring for null metadata. In addition, the dataflow engine tracks invalid
metdata flows in order to enable partial monitoring without false positives.
Our experimental results show that filtering can greatly reduce the average
overheads of monitoring from 7x, 18x, and 11x for UMC, BC, and DIFT down to 4x
for UMC and BC and just 13\% for DIFT. With partial monitoring, we see that
with a 50\% overhead budget, BC can still achieve 83\% coverage and UMC can
still achieve 51\% coverage. Finally, we show that although source dropping can
achieve better coverage than unrestricted dropping when it works, it is not
able to match overhead targets as well as unrestricted dropping.

>>>>

Parallel run-time monitoring techniques are attractive solutions for improving
the reliability, security, and debugging capabilities of systems. In this
paper, we have presented an architecture that reduces the overheads of
monitoring through filtering and adjustable partial monitoring. This is done by
using a hardware dataflow tracking engine in order to track and filter out
monitoring for null metadata. In addition, the dataflow engine tracks invalid
metdata flows in order to enable partial monitoring without false positives.
Our experimental results show that filtering can greatly reduce the average
overheads of monitoring from 7x, 18x, and 11x for UMC, BC, and DIFT down to 4x
for UMC and BC and just 13\% for DIFT. With partial monitoring, we see that
with a 50\% overhead budget, BC can still achieve 83\% coverage and UMC can
still achieve 51\% coverage. Finally, we show that although source-only dropping can
achieve better coverage than unrestricted dropping when it works, it is not
able to match overhead targets as well as unrestricted dropping.

### 2014-05-30 evaluation.tex
Added conclusion.

In this section we evaluate the trade-offs of the three dropping policies
presented in Section~\ref{sec:policies.events}: unrestricted dropping, source
dropping, and sub-flow dropping.

@@

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to enable partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high or total coverage across all runs. Varying the monitoring that is
done for different runs can be achieved by introducing some randomness into the
dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running with a 20\%
overhead budget and an unrestricted dropping policy. Here, when events are
dropped probabilistically with the probability scaled proportionally to slack.
That is, when slack is high, the probability of dropping is low and the
probability increases as we run out of slack. We see that the coverage achieved
quickly asymptotes for an unrestricted dropping policy. This is due to the fact
that since there are multiple dependent monitoring operations that must be done
in order for a final check to be successful, a small amount of randomness at
each event quickly compounds to reduce the probability of certain checks to
near zero. Instead, source dropping is much better suited for achieve high
coverage over multiple runs.

>>>>

In this section we evaluate the trade-offs of the three dropping policies
presented in Section~\ref{sec:policies.events}: unrestricted dropping, source-only
dropping, and sub-flow dropping.

@@

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to enable partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high or total coverage across all runs. Varying the monitoring that is
done for different runs can be achieved by introducing some randomness into the
dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running with a 20\%
overhead budget and an unrestricted dropping policy. Here, when events are
dropped probabilistically with the probability scaled proportionally to slack.
That is, when slack is high, the probability of dropping is low and the
probability increases as we run out of slack. We see that the coverage achieved
quickly asymptotes for an unrestricted dropping policy. This is due to the fact
that since there are multiple dependent monitoring operations that must be done
in order for a final check to be successful, a small amount of randomness at
each event quickly compounds to reduce the probability of certain checks to
near zero. Instead, source-only dropping is much better suited for achieve high
coverage over multiple runs.

### 2014-05-30 policies.tex
More edits based on Ed's comments.
Added conclusion.

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented.

@@

Another possible method is to make the dropping decisions somewhere in between
source-only dropping and unrestricted dropping. From the graph in 
Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
monitoring on that flow. Similarly, event {\tt F} creates a decision point for
the bottom flow that will result in no wasted work and dropping event {\tt A}
allows the entire shown metadata flow to be dropped with no wasted work. If the
program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
made for this policy.
Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there still exists
edge cases where wasted work can occur, but typically we expect much less
wasted work than unrestricted 
dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
certain overhead to fall between source and unrestricted dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target.
Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
and ability to match overhead.

>>>>

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented. This can be difficult to determine precisely due to difficult
to measure overheads such as contention for shared memory. However, we have
found that using only the stalls due to FIFO back pressure works well in
practice.

@@

Another possible method is to make the dropping decisions somewhere in between
source-only dropping and unrestricted dropping. From the graph in 
Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
monitoring on that flow. Similarly, event {\tt F} creates a decision point for
the bottom flow that will result in no wasted work and dropping event {\tt A}
allows the entire shown metadata flow to be dropped with no wasted work. If the
program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
made for this policy.
Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
finer-grained that source-only dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there still exists
edge cases where wasted work can occur, but typically we expect much less
wasted work than unrestricted 
dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
certain overhead to fall between source-only and unrestricted dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target.
Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
and ability to match overhead.
